### 2016-08-05 ms/ms-common-defs.tex
Allocation model and figures

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand{\plant}{\texttt{plant}}

\title{Growth trajectories: How functional traits influence plant growth}
\runninghead{O\lowercase{n the influence of traits on growth}}

\author{Daniel S. Falster\textsuperscript{\textasteriskcentered} \& Richard G. FitzJohn}
\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia
\\
\textsuperscript{\textasteriskcentered} Correspondence author. E-mail: \texttt{adaptive.plant@gmail.com}
}

\usepackage{natbib}
\bibliographystyle{ecol_let}
\setcitestyle{authoryear,open={(},close={)}}
% allows for numbered referencing, as required by ecol letters
\usepackage{etoolbox}
\newbool{MyRefNumbers}
\booltrue{MyRefNumbers} % comment to remove numbers in reference list

\usepackage{graphicx}
\graphicspath{{../output/}}

% We will generate all images so they have a max width and height. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=15cm,height=21cm,keepaspectratio]{#1}}