### 2016-09-20 ms/ms-common-defs.tex
Full draft
- Add abstract, conclusions, minor edits throughout
- Improve figure and table captions
- Add Remko as coauthor
- Add tables in SuppMat, details about hyper-parameterisation
- Better table of contents in SuppMat

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand{\plant}{\texttt{plant}}
\newcommand{\wplcp}{\textsc{wplcp}}
\newcommand{\sma}{\textsc{sma}}

\title{Growth trajectories: how functional traits influence plant growth}
\runninghead{H\lowercase{ow traits influence growth}}

\author{Daniel S. Falster\textsuperscript{1,\textasteriskcentered}, Remko Duursma\textsuperscript{2}, \& Richard G. FitzJohn\textsuperscript{1,3}}
\affiliation{\textsuperscript{1} Biological Sciences, Macquarie University NSW 2109, Australia \\
\textsuperscript{2} Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia \\
\textsuperscript{3} Imperial College, London, United Kingdom \\
\textsuperscript{\textasteriskcentered} Correspondence author. E-mail: \email{adaptive.plant@gmail.com}
}

\usepackage{natbib}
\bibliographystyle{ecol_let}
\setcitestyle{authoryear,open={(},close={)}}
% allows for numbered referencing, as required by ecol letters
\usepackage{etoolbox}
\newbool{MyRefNumbers}
\booltrue{MyRefNumbers} % comment to remove numbers in reference list

\pdfminorversion=4 % tell pdflatex to generate PDF in version 1.4
% Needed for problem-free conversion on manuscript central (arrgh)
% http://timotheepoisot.fr/2013/06/24/manuscript-central-pdf-fix/

\usepackage{graphicx}
\graphicspath{{../output/}}

% We will generate all images so they have a max width and height. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=15cm,height=21cm,keepaspectratio]{#1}}