### 2018-04-05 ms/ms-common-defs.tex
Revise paper after reviewers comments

% abbreviations
\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand{\plant}{\texttt{plant}}
\newcommand{\wplcp}{\textsc{wplcp}}
\newcommand{\lma}{\textsc{lma}}
\newcommand{\wood}{\textsc{wd}}
\newcommand{\seed}{\textsc{sm}}
\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
\newcommand{\nitrogen}{\textsc{n}}
\newcommand{\sma}{\textsc{sma}}

\usepackage{graphicx}
\graphicspath{{../output/}}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Daniel S. Falster}
\author[b]{Remko A. Duursma}
\author[a,c]{Richard G. FitzJohn}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[c]{Imperial College, London, United Kingdom}

% Please give the surname of the lead author for the running footer
\leadauthor{Falster}