### 2018-04-15 ms/ms-supp.Rnw
Minor modifications for submission

\documentclass[9pt,onecolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle}

\title{
{\normalfont \fontsize{12pt}{16pt}
{\color{black50} Supporting information for:}\\} \vskip6pt
\fontsize{16pt}{20pt} How functional traits influence plant growth and shade tolerance across the life-cycle
}

% abbreviations
\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand{\plant}{\texttt{plant}}
\newcommand{\wplcp}{\textsc{wplcp}}
\newcommand{\lma}{\textsc{lma}}
\newcommand{\wood}{\textsc{wd}}
\newcommand{\seed}{\textsc{sm}}
\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
\newcommand{\nitrogen}{\textsc{n}}
\newcommand{\sma}{\textsc{sma}}

\usepackage{graphicx}
\graphicspath{{../output/SI/}}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Daniel S. Falster}
\author[b]{Remko A. Duursma}
\author[c]{Richard G. FitzJohn}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[c]{Imperial College, London, United Kingdom}

% Please give the surname of the lead author for the running footer
\leadauthor{Falster}

\dates{This supplement was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\hypersetup{
     linkcolor=pnasbluetext
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\maketitle

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thesection}{S\arabic{section}}

\section{Additional details on the \texttt{FF16} growth model in the {\plant} package}

In this paper we use the \texttt{FF16} growth model from the package {\plant} \citep{Falster-2016}, which is accessed via \textsc{r} \citep{R-2015}. A full derivation of  Eqns 1-21 from the main text are described is provided in Appendix of \citep{Falster-2016}, see also \citep{Falster-2011}. Below we repeat some additional details from \citep{Falster-2016}, so that readers have a complete picture of the model.

The \texttt{FF16} physiological model includes default values for all needed parameters (Table \ref{tab:params_core}). Species are known to vary considerably in many of these parameters, such as $\phi$, $\rho$, $\nu$, and $\omega$; so by varying parameters one can account for species differences. When altering a parameter in the model, however, one must also consider whether there are trade-offs linking parameters. This is achieved via a hyper-parameterisation function.

\subsection{Hyper-parameterisation of the \texttt{FF16} model via traits}

{\plant} allows for the hyper-parameterisation of the \texttt{FF16} physiological model via plant functional traits: this enables simultaneous variation in multiple parameters in accordance with an assumed trade-off. The hyper-parameterisation function used in the current analysis differs slightly from that in \citep{Falster-2016}. The functions implemented are as follows.

\subsubsection{Seed mass}

Effects of the trait seed mass, are naturally embedded in the equation determining fecundity and the initial height of seedlings. In addition, we let the accessory cost per seed be a multiple of seed size,
$\alpha_{{\rm f3}} = \beta_{{\rm f1}} \omega,$
as empirically observed \citep{Henery-2001}.

\subsubsection{Nitrogen per unit leaf area}

Photosynthesis per unit leaf area and respiration rates per unit leaf mass (or area) are assumed to vary with leaf nitrogen per unit area, $\nu$. The calculation of respiration rates is already described above. To calculate the average annual photosynthesis for a leaf, we integrate the instantaneous rate per unit leaf area over the annual solar trajectory, using a rectangular-hyperbolic photosynthesis light response curve,
\begin{equation} \label{eq:pts}
p(\nu, E) = \frac1{365 \textrm{d}}\int_0^{365 \textrm{d}}\frac{Y(t) + A_{\rm max} - \sqrt{(Y(t) + A_{\rm max})^2 - 4 \beta_{{\rm lf2}} Y(t) A_{\rm max} }}{2 \beta_{{\rm lf2}}} {\rm d}t,
\end{equation}
where $A_{\rm max}$ is the maximum photosynthetic capacity of the leaf, 
$\beta_{{\rm lf2}}$ is the curvature of the light response curve,
$Y(t) = \beta_{{\rm lf3}} I(t)$ is the initial yield of the light response curve, with $\beta_{{\rm lf3}}$ being the quantum yield parameter,
$I(t) = k_{\rm I} \, I_0(t)\, E$ is the intensity of light on the leaf surface, and
$I_0(t)$ is light incident on a surface perpendicular to the sun's rays directly above the canopy at time $t$.
The profile of $I_0(t)$ is given by a solar model adapted from \citep{TerSteege-1997}.

We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as
\begin{equation} \label{eq:amax}
A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}},
\end{equation}
where $\beta_{{\rm lf1}}$ and $\beta_{{\rm lf5}}$ are constants. The relationships is normalized around $\nu_0$, the global mean of leaf nitrogen per unit area.

\subsubsection{Leaf mass per unit area}

The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citep{Wright-2004} (Fig. \ref{fig:leafturn}),
\begin{equation}
k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.
\end{equation}
This relationship is normalised around $\phi_0$, the global mean of $\phi$. This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.

We also vary the mass-based leaf respiration rate so that it stays constant per unit leaf area and varies with $\phi$ and nitrogen per unit leaf area $\nu$, as empirically observed by \citep{Wright-2004},
\begin{equation}
r_{\rm l} = \frac{\beta_{{\rm lf4}}\, \nu}{\phi}.
\end{equation}

\subsubsection{Wood density}

The trait wood density, denoted by $\rho$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. As for $\phi$, these relationships are normalized around $\rho_0$, the global mean of $\rho$. By default, $\beta_{{\rm kI2}}$ and $\beta_{{\rm ks2}}$ are set to zero, so these linkages only become present when these parameters are set to something other than their default values.

The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as
\begin{equation}
r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},
\end{equation}
where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood. Similarly, the rate of bark respiration per unit mass varies as
\begin{equation}
r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},
\end{equation}
with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.

\subsubsection{Implementation of hyper-parameterisation function}

The hyper-parameterisation function used in this analysis is implemented as follows:

<<echo=FALSE, highlight=TRUE, comment = ''>>=
make_hyperpar2
#txt <- capture.output(make_hyperpar2)
#writeLines(strwrap(txt, width = 150, exdent = 10))
@

\section{References}
\bibliography{references}

\section{Supporting Tables and Figures}\label{app:supp_info_table}

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Apr  5 13:08:59 2018
\begin{table*}[ht]
\centering
\caption{Core parameter of the 	exttt{FF16} physiological model.} 
\label{tab:params_core}
\begin{tabular}{p{7cm}llll}
  \hline
Description & Symbol & Unit & Code & Value \\ 
  \hline
\textbf{Plant construction} &  &  &  &  \\ 
  Crown-shape parameter & $\eta$ &  & \texttt{eta} & 12 \\ 
  Leaf mass per area & $\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{lma} & 0.1978791 \\ 
  Wood density & $\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & \texttt{rho} & 608 \\ 
  Sapwood area per unit leaf area & $\theta$ &  & \texttt{theta} & 0.0002141786 \\ 
  Height of plant with leaf area of 1m$^2$ & $\alpha_{\rm l1}$ & m & \texttt{a\_l1} & 2.17 \\ 
  Exponent of relationship between height and leaf area & $\alpha_{\rm l2}$ &  & \texttt{a\_l2} & 0.5 \\ 
  Root mass per unit leaf area & $\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{a\_r1} & 0.07 \\ 
  Ratio of bark area to sapwood area & $\alpha_{\rm b1}$ &  & \texttt{a\_b1} & 0.17 \\ 
  \textbf{Production} &  &  &  &  \\ 
  Leaf photosynthesis per area & $\alpha_{\rm p1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-2}$ & \texttt{a\_p1} & 151.1778 \\ 
  Saturation of leaf photosynthesis per area & $\alpha_{\rm p2}$ &  & \texttt{a\_p2} & 0.2047162 \\ 
  Yield = fraction of carbon fixed converted into mass & $\alpha_{\rm y}$ &  & \texttt{a\_y} & 0.7 \\ 
  Biomass per mol carbon & $\alpha_{\rm bio}$ & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & \texttt{a\_bio} & 0.0245 \\ 
  Leaf respiration per mass & $r_{\rm l}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_l} & 198.4545 \\ 
  Fine-root respiration per mass & $r_{\rm r}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_r} & 217 \\ 
  Sapwood respiration per mass & $r_{\rm s}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_s} & 6.598684 \\ 
  Bark respiration per mass & $r_{\rm b}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_b} & 13.19737 \\ 
  Turnover rate for leaves & $k_{\rm l}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_l} & 0.1521952 \\ 
  Turnover rate for sapwood & $k_{\rm s}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_s} & 0.2 \\ 
  Turnover rate for bark & $k_{\rm b}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_b} & 0.2 \\ 
  Turnover rate for fine roots & $k_{\rm r}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_r} & 1 \\ 
  \textbf{Fecundity} &  &  &  &  \\ 
  Seed mass & $\omega$ & kg & \texttt{omega} & 0.000038 \\ 
  Height at maturation & $H_{{\rm mat}}$ & m & \texttt{hmat} & 15 \\ 
  Maximum allocation to reproduction & $\alpha_{\rm f1}$ &  & \texttt{a\_f1} & 0.8 \\ 
  Parameter determining rate of change in $r(x, m_\textrm{l})$ around $H_{\rm mat}$ & $\alpha_{\rm f2}$ &  & \texttt{a\_f2} & 10 \\ 
   \hline
\end{tabular}
\end{table*}

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Apr  5 13:08:59 2018
\begin{table*}[ht]
\centering
\caption{Parameters for hyper-parameterisation of the \texttt{FF16} physiological model.} 
\label{tab:params_hyper}
\begin{tabular}{p{9cm}llll}
  \hline
Description & Symbol & Unit & Code & Value \\ 
  \hline
\textbf{Leaf turnover} &  &  &  &  \\ 
  Global average leaf mass per area & $\phi_0$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{lma\_0} & 0.1978791 \\ 
  Rate of leaf turnover at average leaf mass per unit leaf area, $\phi_0$ & $\beta_{\rm kl1}$ & $\mathrm{yr}^{-1}$ & \texttt{B\_kl1} & 0.4565855 \\ 
  Scaling exponent for $\phi$ in leaf turnover & $\beta_{\rm kl2}$ &  & \texttt{B\_kl2} & 1.71 \\ 
  \textbf{Sapwood turnover} &  &  &  &  \\ 
  Global average wood density & $\rho_0$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & \texttt{rho\_0} & 608 \\ 
  Rate of sapwood turnover at average wood density, $\rho_0$ & $\beta_{\rm ks1}$ & $\mathrm{yr}^{-1}$ & \texttt{B\_ks1} & 0.2 \\ 
  Scaling exponent for $\rho$ in sapwood turnover & $\beta_{\rm ks2}$ &  & \texttt{B\_ks2} & 1.25 \\ 
  \textbf{Photosynthesis} &  &  &  &  \\ 
  Leaf nitrogen per unit leaf area & $\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{narea} & 0.00187 \\ 
  Global average nitrogen per unit leaf area & $\nu_0$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{narea\_0} & 0.00187 \\ 
  Potential CO$_2$ photosynthesis at average leaf nitrogen, $\nu_0$ & $\beta_{\rm lf1}$ & $\mathrm{mol}\,\mathrm{d}^{-1}\,\mathrm{m}^{-2}$ & \texttt{B\_lf1} & 0.8273474 \\ 
  Curvature of light response curve & $\beta_{\rm lf2}$ &  & \texttt{B\_lf2} & 0.75 \\ 
  Quantum yield of leaf photosynthesis (CO$_2$ per unit photosynthetically active radiation) & $\beta_{\rm lf3}$ &  & \texttt{B\_lf3} & 0.04 \\ 
  Scaling exponent for leaf nitrogen in maximum leaf photosynthesis & $\beta_{\rm lf5}$ &  & \texttt{B\_lf5} & 0.5 \\ 
  \textbf{Respiration} &  &  &  &  \\ 
  CO$_2$ respiration per unit leaf nitrogen & $\beta_{\rm lf4}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{B\_lf4} & 15750 \\ 
  CO$_2$ respiration per unit sapwood volume & $\beta_{\rm rs1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-3}$ & \texttt{B\_rs1} & 4012 \\ 
  CO$_2$respiration per unit bark volume & $\beta_{\rm rb1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-3}$ & \texttt{B\_rb1} & 8024 \\ 
   \hline
\end{tabular}
\end{table*}

\newpage
\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{lma_tradeoff.pdf}
\caption{\textbf{Leaf turnover decreases with leaf-mass per unit leaf area.}
Data from \citep{Wright-2004} for 678 species from 51 sites, each
point giving a species-average. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship.}
\label{fig:leafturn}
\end{figure*}
\newpage

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{assumptions.pdf}
\caption{\textbf{Assumptions of a functional balance model for plant construction.}
Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
Blue lines show standardised major axis lines fit to different species. The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table 4 from main text.}
\label{fig:assumptions}
\end{figure*}

\newpage

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{growth_light_dia.pdf}
\caption{\textbf{Effect of four traits on stem-diameter growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired stem-diameter and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}
\label{fig:growth_light_dia}
\end{figure*}

\newpage

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{growth_light_area.pdf}
\caption{\textbf{Effect of four traits on stem-area growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired stem-area and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}
\label{fig:growth_light_area}
\end{figure*}

\newpage

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{growth_light_mass.pdf}
\caption{\textbf{Effect of four traits on plant mass growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired mass and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}
\label{fig:growth_light_mass}
\end{figure*}

\end{document}