### 2018-05-02 ms/Table2.tex

\documentclass[9pt,onecolumn]{pnas-new}

\newcommand{\plant}{\texttt{plant}}
\newcommand{\wplcp}{\textsc{wplcp}}
\newcommand{\lma}{\textsc{lma}}
\newcommand{\wood}{\textsc{wd}}
\newcommand{\seed}{\textsc{sm}}
\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
\newcommand{\nitrogen}{\textsc{n}}

\begin{document}

\begin{table}[!ht]
 \caption{Key trade-offs (benefit and cost) for the five traits considered, as encoded into the {\plant} model.
 }
\centering
  \begin{tabular}{p{0.5cm}p{3.6cm}p{2.5cm}p{0.4cm}}
  \toprule
  Trait &  Benefit & Cost & Ref. \\
  \midrule
  {\seed} & $\uparrow$ size of seedlings & $\downarrow$ seed production rate & (39)\\
  {\hmat}  & $\uparrow$ growth rate & $\downarrow$ seed production rate & (40)\\
  {\nitrogen} & $\uparrow$ photosynthetic rate in high light & $\uparrow$ respiration rate & (2)\\
  {\lma} & $\downarrow$ leaf turnover rate & $\uparrow$ cost of building leaf & (2)\\
  {\wood} & $\downarrow$ sapwood turnover rate & $\uparrow$ cost of building stem & (2)\\
  \bottomrule
  \end{tabular}

\addtabletext{Traits: {\seed} = seed mass, {\hmat}= height at maturation, {\nitrogen}= leaf nitrogen content per unit leaf area, {\lma}=leaf mass per unit leaf area, {\wood}=wood density.

}
\label{tab:traits}
\end{table}

\end{document}